---
title: "Getting Started with greenAlgoR"
author: "Adrien Taudière"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with greenAlgoR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../pkgdown/assets/bibliography.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

## Introduction

The `greenAlgoR` package provides tools to estimate the carbon footprint and energy consumption of computational tasks in R. This package is based on the Green Algorithms framework [@lannelongue_green_2021], which provides a standardized approach to quantifying the environmental impact of computational research.

Understanding the carbon footprint of our computational work is increasingly important as we strive to make research more sustainable. The `greenAlgoR` package makes it easy to:

- Calculate CO2 emissions from R computations
- Compare different computational approaches
- Optimize code for environmental impact
- Track carbon footprint across research projects

## Installation

```{r, eval=FALSE}
# Install from GitHub (development version)
if (!require("devtools", quietly = TRUE)) {
  install.packages("devtools")
}
devtools::install_github("adrientaudiere/greenAlgoR")
```

```{r setup}
library(greenAlgoR)
library(ggplot2)
```

## Basic Usage

### Calculating Carbon Footprint

The main function `ga_footprint()` calculates the carbon footprint based on several parameters:

```{r basic-example}
# Calculate footprint for a 2-hour computation
result <- ga_footprint(
  runtime_h = 2,
  location_code = "WORLD",  # Global average carbon intensity
  n_cores = 4,
  TDP_per_core = 15,        # Thermal Design Power per core in Watts
  memory_ram = 16           # RAM in GB
)

# View key results
cat("Carbon footprint:", result$carbon_footprint_total_gCO2, "g CO2\n")
cat("Energy needed:", result$energy_needed_kWh, "kWh\n")
```

### Understanding the Results

The function returns a comprehensive list with detailed breakdown:

```{r explore-results}
# View all available information
names(result)

# Key components of carbon footprint
cat("CPU contribution:", result$carbon_footprint_cores, "g CO2\n")
cat("Memory contribution:", result$carbon_footprint_memory, "g CO2\n")
cat("Total footprint:", result$carbon_footprint_total_gCO2, "g CO2\n")
```

## Location-Specific Carbon Intensity

Carbon intensity varies significantly by location due to different energy sources:

```{r location-comparison}
# Compare carbon footprint across different locations
locations <- c("FR", "WORLD", "US", "CN", "NO")
footprints <- sapply(locations, function(loc) {
  ga_footprint(runtime_h = 1, location_code = loc, n_cores = 2)$carbon_footprint_total_gCO2
})

# Create comparison data frame
comparison_df <- data.frame(
  Location = locations,
  CO2_emissions = footprints
)

print(comparison_df)
```

```{r location-plot}
# Visualize the comparison
ggplot(comparison_df, aes(x = reorder(Location, CO2_emissions), y = CO2_emissions)) +
  geom_col(fill = "steelblue", alpha = 0.7) +
  labs(
    title = "Carbon Footprint by Location",
    subtitle = "1-hour computation with 2 cores",
    x = "Location",
    y = "CO2 Emissions (g)",
    caption = "Based on regional carbon intensity differences"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Hardware Configuration

Different hardware configurations have varying environmental impacts:

```{r hardware-comparison}
# Compare different CPU configurations
cpu_configs <- data.frame(
  Config = c("Laptop", "Workstation", "Server"),
  Cores = c(4, 8, 16),
  TDP_per_core = c(10, 15, 25),
  Memory = c(8, 32, 64)
)

# Calculate footprint for each configuration
cpu_configs$Footprint <- mapply(function(cores, tdp, mem) {
  ga_footprint(
    runtime_h = 1,
    n_cores = cores,
    TDP_per_core = tdp,
    memory_ram = mem
  )$carbon_footprint_total_gCO2
}, cpu_configs$Cores, cpu_configs$TDP_per_core, cpu_configs$Memory)

print(cpu_configs)
```

## Current R Session Footprint

You can easily calculate the carbon footprint of your current R session:

```{r session-footprint}
# Get current session footprint
session_fp <- ga_footprint(runtime_h = "session")

cat("Current session footprint:", session_fp$carbon_footprint_total_gCO2, "g CO2\n")
cat("Session runtime:", session_fp$runtime_h, "hours\n")
```

## Visualization with Reference Values

The package includes reference values to put your footprint in context:

```{r reference-visualization, fig.width=8, fig.height=6}
# Calculate footprint with reference values
result_with_ref <- ga_footprint(
  runtime_h = 2,
  n_cores = 4,
  memory_ram = 16,
  add_ref_values = TRUE
)

# Create visualization comparing to reference values
ref_data <- result_with_ref$ref_value
ref_data$is_computation <- FALSE
ref_data$is_computation[ref_data$variable == "Total"] <- TRUE

# Add our computation to the data
computation_data <- data.frame(
  variable = "Your Computation",
  value = result_with_ref$carbon_footprint_total_gCO2,
  prop_footprint = NA,
  is_computation = TRUE
)

plot_data <- rbind(ref_data[, c("variable", "value", "is_computation")], 
                   computation_data[, c("variable", "value", "is_computation")])
plot_data$value <- as.numeric(plot_data$value)

ggplot(plot_data, aes(x = reorder(variable, value), y = value, 
                      fill = is_computation)) +
  geom_col(alpha = 0.8) +
  scale_fill_manual(values = c("FALSE" = "lightblue", "TRUE" = "darkred"), 
                    name = "Type",
                    labels = c("Reference", "Your Computation")) +
  scale_y_log10() +
  coord_flip() +
  labs(
    title = "Carbon Footprint Comparison",
    subtitle = "Your computation vs. reference activities",
    x = "Activity",
    y = "CO2 Emissions (g, log scale)",
    caption = "Reference values help contextualize computational impact"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Best Practices

1. **Choose efficient algorithms**: Optimize your code to reduce runtime
2. **Consider location**: Run computations in regions with cleaner energy
3. **Right-size resources**: Use appropriate CPU/memory for your task
4. **Monitor regularly**: Track footprint across projects
5. **Share awareness**: Include carbon footprint in research reporting

## Next Steps

- Explore the `ga_targets()` function for pipeline analysis
- Check the package documentation for advanced configuration options
- Consider the carbon impact in your research workflow decisions

## References