---
title: "Using greenAlgoR with targets Pipelines"
author: "Adrien Taudière"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using greenAlgoR with targets Pipelines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../pkgdown/assets/bibliography.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/targets-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
```

## Introduction

The `targets` package [@landau_targets_2021] provides a powerful framework for reproducible computational workflows in R. The `greenAlgoR` package seamlessly integrates with `targets` to help you understand the environmental impact of your entire data analysis pipeline.

This vignette demonstrates how to:

- Calculate carbon footprint for complete `targets` pipelines
- Identify the most carbon-intensive steps in your workflow
- Optimize pipelines for environmental efficiency

## Setup

```{r setup}
library(greenAlgoR)
library(targets)
library(ggplot2)
```

## Basic targets Integration

The `ga_targets()` function analyzes your `targets` pipeline and calculates the total carbon footprint based on:

- Runtime of each target
- Memory usage patterns
- Storage requirements (optional)

### Simple Example

```{r targets-example}
# Create a simple targets example
tar_dir({ # tar_dir() runs code from a temp dir for CRAN compatibility
  
  # Define a simple pipeline
  tar_script({
    library(targets)
    
    list(
      tar_target(
        name = data_prep,
        command = {
          # Simulate data preparation (2 seconds)
          Sys.sleep(2)
          data.frame(x = rnorm(1000), y = rnorm(1000))
        }
      ),
      tar_target(
        name = analysis,
        command = {
          # Simulate analysis (1 second)
          Sys.sleep(1)
          lm(y ~ x, data = data_prep)
        }
      ),
      tar_target(
        name = visualization,
        command = {
          # Simulate plotting (0.5 seconds)
          Sys.sleep(0.5)
          plot(data_prep$x, data_prep$y)
          "plot_completed"
        }
      )
    )
  }, ask = FALSE)
  
  # Run the pipeline
  tar_make()
  
  # Get metadata
  metadata <- tar_meta()
  print(metadata[, c("name", "seconds", "bytes")])
  
  # Calculate carbon footprint
  pipeline_footprint <- ga_targets(
    tar_meta_raw = metadata,
    location_code = "WORLD",
    n_cores = 2,
    TDP_per_core = 15,
    memory_ram = 8
  )
  
  cat("Pipeline carbon footprint:", 
      pipeline_footprint$carbon_footprint_total_gCO2, "g CO2\n")
  cat("Total runtime:", pipeline_footprint$runtime_h * 3600, "seconds\n")
})
```

## Advanced Pipeline Analysis

For more complex pipelines, you can get detailed insights:

```{r advanced-targets}
tar_dir({
  # Create a more complex pipeline
  tar_script({
    library(targets)
    
    simulate_computation <- function(duration, size = 1000) {
      Sys.sleep(duration)
      matrix(rnorm(size * size), nrow = size)
    }
    
    list(
      tar_target(small_task, simulate_computation(0.5, 100)),
      tar_target(medium_task, simulate_computation(2, 500)),
      tar_target(large_task, simulate_computation(5, 1000)),
      tar_target(
        combined_analysis, 
        {
          # Combine results
          result <- list(
            small = summary(small_task),
            medium = summary(medium_task), 
            large = summary(large_task)
          )
          Sys.sleep(1)  # Additional processing time
          result
        }
      )
    )
  }, ask = FALSE)
  
  tar_make()
  metadata <- tar_meta()
  
  # Calculate footprint with storage estimation
  detailed_footprint <- ga_targets(
    tar_meta_raw = metadata,
    location_code = "FR",  # France has relatively low carbon intensity
    n_cores = 4,
    TDP_per_core = 20,
    memory_ram = 16,
    add_storage_estimation = TRUE
  )
  
  # Display breakdown
  cat("Total CO2 emissions:", detailed_footprint$carbon_footprint_total_gCO2, "g\n")
  cat("CPU contribution:", detailed_footprint$carbon_footprint_cores, "g\n")
  cat("Memory contribution:", detailed_footprint$carbon_footprint_memory, "g\n")
  
  if (!is.null(detailed_footprint$power_draw_storage_kWh)) {
    storage_co2 <- detailed_footprint$carbon_intensity * detailed_footprint$power_draw_storage_kWh
    cat("Storage contribution:", storage_co2, "g\n")
  }
})
```

## Comparing Different Configurations

You can compare how different hardware configurations affect your pipeline's carbon footprint:

```{r config-comparison}
tar_dir({
  # Use the same pipeline as above
  tar_script({
    library(targets)
    
    simulate_computation <- function(duration, size = 1000) {
      Sys.sleep(duration)
      matrix(rnorm(size * size), nrow = size)
    }
    
    list(
      tar_target(task1, simulate_computation(1, 200)),
      tar_target(task2, simulate_computation(2, 300)),
      tar_target(task3, simulate_computation(1.5, 250))
    )
  }, ask = FALSE)
  
  tar_make()
  metadata <- tar_meta()
  
  # Compare different configurations
  configs <- data.frame(
    Config = c("Laptop", "Desktop", "Server"),
    Cores = c(2, 8, 16),
    TDP = c(10, 15, 25),
    RAM = c(8, 16, 64),
    Location = c("WORLD", "FR", "NO")  # Different locations
  )
  
  # Calculate footprint for each configuration
  configs$CO2_emissions <- mapply(function(cores, tdp, ram, loc) {
    ga_targets(
      tar_meta_raw = metadata,
      n_cores = cores,
      TDP_per_core = tdp,
      memory_ram = ram,
      location_code = loc
    )$carbon_footprint_total_gCO2
  }, configs$Cores, configs$TDP, configs$RAM, configs$Location)
  
  print(configs)
})
```

## Visualizing Pipeline Carbon Footprint

Create comprehensive visualizations of your pipeline's environmental impact:

```{r pipeline-visualization, fig.width=10, fig.height=8}
tar_dir({
  # Create pipeline and get footprint with reference values
  tar_script({
    library(targets)
    list(
      tar_target(data_load, {Sys.sleep(1); rnorm(5000)}),
      tar_target(preprocessing, {Sys.sleep(3); scale(data_load)}),
      tar_target(modeling, {Sys.sleep(5); lm(data_load ~ seq_along(data_load))}),
      tar_target(postprocessing, {Sys.sleep(2); summary(modeling)})
    )
  }, ask = FALSE)
  
  tar_make()
  metadata <- tar_meta()
  
  pipeline_result <- ga_targets(
    tar_meta_raw = metadata,
    location_code = "WORLD",
    n_cores = 4,
    memory_ram = 16,
    add_ref_values = TRUE
  )
  
  # Create comprehensive visualization
  ref_data <- pipeline_result$ref_value
  ref_data$category <- "Reference Activities"
  
  # Add pipeline data
  pipeline_data <- data.frame(
    variable = "Your Pipeline",
    value = pipeline_result$carbon_footprint_total_gCO2,
    prop_footprint = NA,
    category = "Your Computation"
  )
  
  # Combine data
  plot_data <- rbind(
    ref_data[, c("variable", "value", "category")],
    pipeline_data[, c("variable", "value", "category")]
  )
  plot_data$value <- as.numeric(plot_data$value)
  
  # Create the plot
  ggplot(plot_data, aes(x = reorder(variable, value), y = value, 
                        fill = category)) +
    geom_col(alpha = 0.8) +
    scale_fill_manual(
      values = c("Reference Activities" = "lightblue", 
                "Your Computation" = "darkgreen"),
      name = "Type"
    ) +
    scale_y_log10() +
    coord_flip() +
    labs(
      title = "Carbon Footprint of targets Pipeline",
      subtitle = paste("Total emissions:", 
                       round(pipeline_result$carbon_footprint_total_gCO2, 2), 
                       "g CO2"),
      x = "Activity",
      y = "CO2 Emissions (g, log scale)",
      caption = "Comparison with everyday reference activities"
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom",
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 12)
    )
})
```

## Target-Level Analysis

For detailed optimization, you might want to analyze individual targets:

```{r target-analysis}
tar_dir({
  tar_script({
    library(targets)
    list(
      tar_target(quick_task, {Sys.sleep(0.5); "done"}),
      tar_target(slow_task, {Sys.sleep(10); "done"}),
      tar_target(memory_intensive, {
        # Simulate memory-intensive task
        big_matrix <- matrix(rnorm(1000*1000), nrow = 1000)
        Sys.sleep(3)
        summary(big_matrix)
      })
    )
  }, ask = FALSE)
  
  tar_make()
  metadata <- tar_meta()
  
  # Analyze each target separately
  target_analysis <- data.frame(
    Target = metadata$name,
    Runtime_sec = metadata$seconds,
    Memory_MB = metadata$bytes / (1024^2),
    stringsAsFactors = FALSE
  )
  
  # Calculate individual footprints (simplified)
  target_analysis$CO2_estimate <- sapply(metadata$seconds, function(sec) {
    ga_footprint(
      runtime_h = sec / 3600,
      n_cores = 2,
      memory_ram = 8
    )$carbon_footprint_total_gCO2
  })
  
  print(target_analysis)
  
  # Identify the most carbon-intensive target
  most_intensive <- target_analysis[which.max(target_analysis$CO2_estimate), ]
  cat("\nMost carbon-intensive target:", most_intensive$Target, 
      "(", round(most_intensive$CO2_estimate, 3), "g CO2 )\n")
})
```

## Best Practices for Sustainable Pipelines

1. **Profile your pipeline**: Use `tar_meta()` to identify bottlenecks
2. **Optimize slow targets**: Focus on reducing runtime of carbon-intensive steps
3. **Cache efficiently**: Use `targets` caching to avoid re-running expensive computations
4. **Choose appropriate hardware**: Match computational resources to task requirements
5. **Consider location**: Run pipelines in regions with cleaner energy if possible

## Integration with Workflow

Include carbon footprint reporting as part of your standard pipeline:

```{r workflow-integration, eval=FALSE}
# Add to your _targets.R file
tar_target(
  pipeline_footprint,
  {
    # Calculate footprint after pipeline completion
    footprint <- ga_targets(location_code = "FR")
    
    # Log the results
    cat("Pipeline carbon footprint:", 
        footprint$carbon_footprint_total_gCO2, "g CO2\n")
    
    # Save for reporting
    saveRDS(footprint, "results/carbon_footprint.rds")
    
    footprint
  }
)
```

## Conclusion

Integrating `greenAlgoR` with `targets` provides a powerful way to understand and optimize the environmental impact of your computational workflows. By regularly monitoring carbon footprint, you can make informed decisions about computational efficiency and contribute to more sustainable research practices.

## References